### 2021-06-05 Paper/Paper_Draft.Rmd
060421 update

---
title: 'Predicting Enrollment with Stacked Models: Bringing Together Theory and Empirics'
author: "Haroon Atcha"
date: "6/4/2021"
output: html_document

---

# Introduction

# Data

## Simulating Data

The data used in this paper was generated specifically for the purpose and was **not** gathered at any specific community college. This was done for several reasons:

1. Simulated data are generated with *known* parameters, allowing us to determine how accurately models recover *true* estimates
2. Simulated data allow for greater generalizability. Data gathered at one institution may have been generated by a process that differs dramatically from other contexts
3. Simulated data allow us to demonstrate theoretically important trends with  because we *know* all relevant parameters

## Data Generating Process

The data used in this study was generated through two processes. First, the number of new enrollees in a given semester was generated at the aggregate level (i.e. total number of new students). Second, the retention status of enrolled students in a given semester was generated at the individual level (i.e. did student n return at t + 1). Finally, the individual level predictions were aggregated and summed with the number of new enrollees to produce total headcount for a given semester. This process is visualized in chart 1.

```{r, echo = FALSE}
knitr::include_graphics('Data Generation Process.png')
```

### Returning Enrollees

Existing students are known at the individual level and have three observed variables: Gender, current semester credit load, and cumulative credits taken. Each student's likelihood of returning in the subsequent semester is a linear function of their gender and curvilinear function of cumulative credits taken such that students are *more* likely to be retained as they approach graduation and *less* likely thereafter. Constant and error terms are also added.

The formula used to generate each student's probability of returning is given below as are the coefficients associated with each term. As noted above, I chose to generate data using a simple model for the purpose of demonstration; Gender and cumulative credit load are widely cited predictors of retention and the values chosen for them fall within reasonable bounds.

logit(y) = $\beta_{Gender}$ + ${\beta_{Cumulative\ Credits}}^2$ + $\beta_{Cumulative\ Credits}$ + *C* + $\epsilon$

```{r, echo = FALSE}
table <- cbind(
  c('Gender', 'Credits', 'Cumulative Credits', 'Likelihood of Return', 'Return'),
  c('Binary', 'Interval', 'Interval', 'Ratio', 'Binary'),
  c(0, 1, 1, 0.02, 0),
  c(1, 21, 121, 0.83, 1),
  c(0.5, 9, 135, 0.72, 0.72),
  c('Sample (0:1)', 'Sample Truncated Normal', '$\\sum_{(i,j) = 1}^n n_{i,j}$', 'Linear Function', 'Sample (0:1)'),
  c('P(1) = 0.5', '$\\mu$ = 6, $\\sigma$ = 9', '-', 'See Equation 1', 'P(1) = Likelihood of Return'))

colnames(table) <- c('Variable', 'Level', 'Min', 'Max', 'Mean', 'Generation', 'Parameters')

knitr::kable(table)

```

### New Enrollees

The number of new enrollees in a given semester was generated at the aggregate level. No individual level characteristics are known about new enrollees *until their first semester*. The number of new enrollees in a given semester is a linear function of change in GDP and semester; values for the former were drawn randomly from a normal distribution with $\mu$ = 0 and $\sigma$ = 1. Constant and error terms are also added with final output of the model scaled up and rounded to an integer value.

The formula used to generate the number of new students each semester is given below and coefficients associated with each term can be found in the Appendix. As with the returning students model, this model was built with an eye towards parsimony and generalizability.

```{r, echo = FALSE}
table <- cbind(
c('$\\beta_{GDP}$', '$\\beta_{Spring\ Semester}$', '$\\beta_{Summer\ Semester}$', '$\\beta_{Fall\ Semester}$',
  '$C_{New Students}$', '$\\varepsilon_{New Students}$', '$\\beta_{Gender}$', '$\\beta_{Cumulative\ Credits}$',
  '$C_{Returning Students}$', '$\\varepsilon_{Returning Students}$'),
c(rep('New', 6), rep('Returning', 4)),
c(2, 3, 0, 6, 0, '$\\ N(\\mu = 0, \\sigma = 1)$', 0.1, 0.02, 0.9, '$\\ N(\\mu = 0, \\sigma = 1)$')
)

colnames(table) <- c('Term', 'Model', 'Value')

knitr::kable(table)

```

## Summary Statistics

# Methods

# Results

# Discussion

In this paper, I have demonstrated the value in bringing theoretical knowledge to bear on predictive models of enrollment. Community colleges construct such models under unique constraints. Moreover, the financial consequences of inaccurate models is more acutely felt than at other types of institutions of higher education. While the extant literature has sought improvements to predictive accuracy by using 'brute-force' methods - e.g. increasing the number of features in models, implementing more model types, exhaustively searching across model hyper-parameters - I argue that these efforts essentially re-invent the wheel insofar as they 'rediscover' relationships that are already well established. Our theoretical knowledge of the predictors of enrollment and persistence is robust. That knowledge should be reflected in our empirical models.

# Future Research

### 2021-06-05 Paper_Outline.Rmd
060421 update

---
title: "Paper Outline"
author: "Haroon Atcha"
date: "6/4/2021"
output: html_document
---

# Summary

Outlines help me write. They are good at organizing my thoughts and help me put citations and figures in the right place. A clean outline makes draft writing much easier. This is the outline for the paper that should hopefully come out of this project.

# Outline

1. Introduction
    - Problems
      - Community Colleges operate under unique data constraints, most of the literature on enrollment management is for selective-enrollment universities
        - Cite: Aksenova, Chen, Nandeshwar, Slim (From selective enrollment universities)
        - cite: Bender, Lawrence, Pennington (from community colleges)
      - Predictive models of enrollment don't reflect our theoretical models
        - Not all processes are ARIMA; cite: Chen, Pfitzner
        - You should model different processes differently; cite: Chen, Trusheim
    - Argument: We can do better by having our statistical models match our theoretical models

2. Data
    - Show Data Generation Chart
    - Data Generation Process
      - Returning students (Linear model)
      - New students (Linear with time series terms)
    - Summary Statistics
      - Prove that the summary statistics are close to what we'd expect in an actual case study
      - Make the case for parsimonious models (since I only used 1-2 variables for each process)
      - Show enrollment time series

3. Methods
    - Aggregate models
      - ARIMA
      - Linear model
    - Stacked model
      - Logit component (returning)
      - ARIMA component (new enrollees)

4. Results
    - Robustness checks
      - Individual level confusion matrix
      - MAPE, MAE, and MSE table for each model type
        - ARIMA only
        - Linear only
        - Stacked linear
        - Stacked curvilinear
    - Toy example has prediction rates close to what case studies show
    - **Even a misspecified stacked model predicts better if it's closer to the theoretical process**

5. Conclusion
    - Predictive models perform better when they are reflective of our theoretical knowledge

### 2021-06-06 Paper/Paper_Draft.Rmd
060521 updates

@@
 date: "6/4/2021"
-output: html_document
-
+output:
+  word_document: default
+  html_document: default
+bibliography: Citations.bib
 ---
@@
 
-# Data
+Enrollment forecasting is one of the most consequential tasks a given Institutional Research department will undertake.
+The pressures of getting it 'right' are especially pronounced in the community college setting.
+As open enrollment institutions that are uniquely reliant on public funding, accurate enrollment forecasts are critical for guaranteeing adequate state-level funding and proper planning.
+Moreover, community colleges operate under unique data availability constraints and admission processes that make the task difficult.
+
+Despite the importance of this task, the extant literature on enrollment forecasting - especially in the community college context - is slim and relatively disconnected from theoretical advances in the study of the predictors of retention and enrollment.
+This manifests itself in two main problems.
+
+First, practitioners operating in open enrollment contexts have little recourse to literature that speaks to their specific circumstances.
+Much of the literature has been generated by practitioners in selective-enrollment institutions and speaks to issues particular to that context.
+[@aksenova_enrollment_2006], [@chen_integrated_2008], [@nandeshwar_enrollment_2009], and [@slim_predicting_2018] for instance, exclusively apply their models to the selective enrollment context.
+Indeed, X's models are built to specifically predict the likelihood of enrollment *after* acceptance; a process that has no direct analogue in the open enrollment context.
+Work addressing the problem in the community college context is largely dated.
+While a highly useful survey of forecasting method used at community colleges, [@bender_community_1981]'s work is now several decades old and understandably is not reflective of modern forecasting tools and approaches.
+Likewise, [@lawrence_forecasting_1980] is useful as a snapshot of best forecasting practices of the time period, but is similarly dated.
+[@pennington_community_2002] represents the most recent and comprehensive project addressing this issue.
+Numerous dissertations (see: x, y, and z) have reproduced and expanded this work in varying local contexts, adding important contextual knowledge but few explicitly addressing modeling choices.
+In all, the literature that speaks to predictive modeling of enrollment in the community college context is limited.
+
+Second, predictive models of enrollment do not reflect our theoretical understanding of the processes that are being modeled.
+Much of the modeling in the extant literature falls into one of two camps: time series approaches at the aggregate level (ARIMA) or linear modeling techniques (OLS, SVG, etc.).
+While such approaches produce perfectly acceptable predictions, such models are rarely reflective of the theoretical processes they attempt to capture.
+[@chen_integrated_2008], for instance, compares predictions from both, ARIMA and linear models though uses the aggregate number of enrollments in a given semester as the target variable despite noting that '[...] the portion of a university’s student enrollees (freshmen) depends on the number of high school graduates within the state [...] [and] the same pool of eligible enrolled or returning students (one-year lagged OSU enrollment).' The recognition that two distinct sub-populations constitute the population of enrollees is not reflected in the modeling which treats each semester's cohort uniformly^[See also [@pfitzner_shortterm_1987] for an early implementation of time series modeling in the community college context that approaches the problem at the aggregate level].
+By contrast, [@trusheim_predictive_2011] does model returning and new enrollees separately, and while this serves as a good example of using theoretical knowledge as a foundation for in empirical exercises, the methods used to achieve predictions are fairly unsophisticated.
+
+Here, I argue that, in the interest of predictive accuracy and knowledge accumulation, our statistical models ought to be theoretically informed.
+To demonstrate this, I present a hypothetical example with generated data, an approach widely used in the study of statistics.
+I show that bringing theoretical knowledge to bear on predictive problems - even if that knowledge is tentative or incomplete - is a fruitful avenue for improving forecast accuracy.
 
@@
 1. Simulated data are generated with *known* parameters, allowing us to determine how accurately models recover *true* estimates
-2. Simulated data allow for greater generalizability. Data gathered at one institution may have been generated by a process that differs dramatically from other contexts
-3. Simulated data allow us to demonstrate theoretically important trends with  because we *know* all relevant parameters
+2. Simulated data allow for generalizability. Data generating processes may differ dramatically across institutions.
+3. Simulated data allow us to demonstrate theoretically important concepts by assuring these concepts are the *only* relevant ones in our model.
 
@@
 
-Existing students are known at the individual level and have three observed variables: Gender, current semester credit load, and cumulative credits taken.
+The retention status of currently enrolled students was generated at the individual level.
+Three variables are observed for each student: Gender, current semester credit load, and cumulative credits taken.
 Each student's likelihood of returning in the subsequent semester is a linear function of their gender and curvilinear function of cumulative credits taken such that students are *more* likely to be retained as they approach graduation and *less* likely thereafter.
-Constant and error terms are also added.
+Constant and error terms are also included.
 
-The formula used to generate each student's probability of returning is given below as are the coefficients associated with each term.
-As noted above, I chose to generate data using a simple model for the purpose of demonstration; Gender and cumulative credit load are widely cited predictors of retention and the values chosen for them fall within reasonable bounds.
+The formula used to generate each student's probability of returning is given below; the coefficients associated with each term as well as the values for the constant and error parameters can be found in the Appendix.
+As noted above, I chose to generate data using a simple model for the purpose of demonstration; Gender and cumulative credit load are widely cited predictors of retention.
 
 logit(y) = $\beta_{Gender}$ + ${\beta_{Cumulative\ Credits}}^2$ + $\beta_{Cumulative\ Credits}$ + *C* + $\epsilon$
+
+### New Enrollees
+
+The number of new enrollees in a given semester was generated at the aggregate level and is given in heuristic form below.
+The number of new enrollees in a given semester is a linear function of change in GDP and semester such that higher GDP growth is associated with greater enrollment.
+Moreover, Fall, Spring, and Summer semester terms each add a constant number of new enrollees in decreasing order.
+Results from the model were then de-normalized and rounded to generate an integer number of new enrollees.
+The coefficients associated with each term as well as the values for the constant and error parameters can be found in the Appendix.
+As with the individual level data, local GDP change and semester effects are widely documented in the literature^[Note that, since GDP data were *not* gathered from actual sources and were instead generated using general sampling process, it could just as easily be interpreted as the percent change in number of graduating high school students.
+For ease of interpretation and generalizabiilty, I choose to present the variable as change in GDP.].
+
+y = $\beta_{GDP}$ + $\beta_{Spring}$ + $\beta_{Summer}$ + $\beta_{Fall}$ + *C* + $\epsilon$
+
+New Students = ($\sigma_{New\ Students}$ * y) + $\mu_{New\ Students}$
+
+# Methods
+
+Modeling choices ought to reflect our underlying knowledge about the data generating process.
+In this example, we know that the processes that generated our observations vary across sub-populations.
+Different generative models determine the number of new enrollees and returning students in a given semester.
+Rather than fitting a single model to the aggregate count of enrollees, which would entail a misalignment between theory and practice and require us to throw away a large amount of information at the individual level, we can estimate a 'stacked' model instead, fitting different models to different sub-populations and aggregating those predictions.
+
+To demonstrate the value in building theory-informed predictive models, I fit and compare several such models.
+I test autoregressive integrated moving average (ARIMA) and simple linear models at the aggregate level, as these are the most common approaches in the extent literature and compare them to three stacked models.
+For the first stacked model, I predict the number of returning students at the individual level using logistic regression and use that number as a linear predictor of the *total* number of enrolling students at semester t + 1.
+Second, I fit a theoretically misspecified model, predicting the number of returning students at the individual level using *only* lower order terms in a logistic regression and the number of new students using ARIMA.
+Finally, I fit a correctly specified model which differs only insofar as it uses a correct higher order term in the individual prediction section.
+
+I compare the results of these models using the Mean Absolute Percent Error (MAPE), Mean Absolute Standard Error (MASE), and Root Mean Squared Error (RMSE), three widely used measures of model accuracy.
+
+# Results
+
+# Discussion
+
+A skeptical reader may justifiably question the value in the above exercise.
+After all, of course the better specified model predicted more accurately, we *knew* the correct parameters before starting.
+This is, however, exactly the point.
+Often, practitioners approach enrollment prediction from a point of imposed ignorance.
+Brute force and automated methods of feature selection are treated as substitutes for theoretically informed modeling choices^[This is not to argue that automated and brute force methods of feature selection are inherently bad.
+Note that the ARIMA model fit in the stacked version above iterates through all combinations of parameters to find the best fit.
+Such methods are powerful tools but they are best leveraged when informed by theoretically grounded modeling choices.].
+In the community college context, where data availability can be limited and the process that generates enrollment data is idiosyncratic in known ways, discarding such knowledge has practical consequences for the accuracy of our enrollment forecasts and consequently, on resource availability.
+Practitioners ought to take pains to incorporate theoretical knowledge in predictive forecasts.
+Employing stacked models, as shown above, can be a simple manner of doing so.
+
+# Future Research
+
+\newpage
+
+# Appendix A: Coefficients, Parameters, and Variable Values
+
+## Coefficients and Model Parameters
+
+```{r, echo = FALSE}
+table <- cbind(
+c('$\\beta_{GDP}$', '$\\beta_{Spring\ Semester}$', '$\\beta_{Summer\ Semester}$', '$\\beta_{Fall\ Semester}$',
+  '$C_{New Students}$', '$\\varepsilon_{New Students}$', '$\\beta_{Gender}$', '$\\beta_{Cumulative\ Credits}$',
+  '$C_{Returning Students}$', '$\\varepsilon_{Returning Students}$'),
+c(rep('New', 6), rep('Returning', 4)),
+c(2, 3, 0, 6, 0, '$\\ N(\\mu = 0, \\sigma = 1)$', 0.1, 0.02, 0.9, '$\\ N(\\mu = 0, \\sigma = 1)$')
+)
+
+colnames(table) <- c('Term', 'Model', 'Value')
+
+knitr::kable(table)
+
+```
+
+## Individual Variable Summary Statistics
 
@@
 
-### New Enrollees
-
-The number of new enrollees in a given semester was generated at the aggregate level.
-No individual level characteristics are known about new enrollees *until their first semester*.
-The number of new enrollees in a given semester is a linear function of change in GDP and semester; values for the former were drawn randomly from a normal distribution with $\mu$ = 0 and $\sigma$ = 1.
-Constant and error terms are also added with final output of the model scaled up and rounded to an integer value.
-
-The formula used to generate the number of new students each semester is given below and coefficients associated with each term can be found in the Appendix.
-As with the returning students model, this model was built with an eye towards parsimony and generalizability.
-
-```{r, echo = FALSE}
-table <- cbind(
-c('$\\beta_{GDP}$', '$\\beta_{Spring\ Semester}$', '$\\beta_{Summer\ Semester}$', '$\\beta_{Fall\ Semester}$',
-  '$C_{New Students}$', '$\\varepsilon_{New Students}$', '$\\beta_{Gender}$', '$\\beta_{Cumulative\ Credits}$',
-  '$C_{Returning Students}$', '$\\varepsilon_{Returning Students}$'),
-c(rep('New', 6), rep('Returning', 4)),
-c(2, 3, 0, 6, 0, '$\\ N(\\mu = 0, \\sigma = 1)$', 0.1, 0.02, 0.9, '$\\ N(\\mu = 0, \\sigma = 1)$')
-)
-
-colnames(table) <- c('Term', 'Model', 'Value')
-
-knitr::kable(table)
-
-```
-
-## Summary Statistics
-
-# Methods
-
-# Results
-
-# Discussion
+# Appendix B: Cut Parts
 
@@
 
-# Future Research
+# Appendix C: Summary Statistics

### 2021-06-06 Data Generation Diagram.R

@@ -1,110 +1,0 @@
-library('DiagrammeR')
-
-grViz("
-digraph data_generation {
-
-  graph[rankdir = TD
-  label = 'Data Generation Process'
-  labelloc = t
-  fontsize = 25]
-
-      subgraph cluster_0{
-      
-      graph[shape = rectangle
-      label = 'Individual-level'
-      labelloc = t
-      fontsize = 25
-      rankdir = TD]
-      
-      node[shape = 'box']
-      b[label = 'Logit@^{1}', shape = 'cylinder']
-      c[label = 'Predict probability of return']
-      d[label = 'Generate return status']
-      e[label = 'Non-returning students']
-      f[label = 'Returning Student list']
-      n[label = 'rnorm@^{3}', shape = 'cylinder']
-      r[label = 'sample 0/1@^{4}', shape = 'cylinder']
-      s[label = '0 Credits']
-      t[label = 'Generate credit load']
-      
-      {rank = same; b, c}
-      {rank = same; e, f}
-      {rank = same; r, d}
-      {rank = same; n, t, s}
-      
-      #edges
-      b -> c
-      d -> {e, f}
-      c -> d
-      r -> d
-      e -> s
-      f -> t
-      n -> t
-      }
-      
-      subgraph cluster_1{
-      graph[shape = rectangle
-      label = 'Aggregate-level'
-      labelloc = t
-      fontsize = 25]
-      
-      node[shape = 'box']
-      g[label = 'Predict # of new students']
-      h[label = 'OLS@^{2}', shape = 'cylinder']
-      i[label = 'Sample credit load']
-      m[label = 'rnorm@^{3}', shape = 'cylinder']
-      o[label = 'Generate individual values']
-      v[label = 'Sample gender + credits@^{5}', shape = 'cylinder']
-      
-      {rank = same; g, h}
-      {rank = same; i, m}
-      {rank = same; v, o}
-      
-      #edges
-      m -> i
-      h -> g
-      g -> o
-      o -> i
-      v -> o
-      }
-      
-      subgraph cluster_2{
-      graph[label = 'Footnotes'
-      labelloc = t
-      penwidth = 0
-      ranksep = 0.02]
-      
-      node[shape = 'none']
-      x[label = '1: logit(y) = \u03b2*Gender + \u03b2*Cumulative credits + c(function of GDP) + \u03b5', x = 1, y = 1]
[…]

### 2021-06-06 Data Generation/Data Generation Diagram.R

@@ -1,0 +1,110 @@
+library('DiagrammeR')
+
+grViz("
+digraph data_generation {
+
+  graph[rankdir = TD
+  label = 'Data Generation Process'
+  labelloc = t
+  fontsize = 25]
+
+      subgraph cluster_0{
+      
+      graph[shape = rectangle
+      label = 'Individual-level'
+      labelloc = t
+      fontsize = 25
+      rankdir = TD]
+      
+      node[shape = 'box']
+      b[label = 'Logit@^{1}', shape = 'cylinder']
+      c[label = 'Predict probability of return']
+      d[label = 'Generate return status']
+      e[label = 'Non-returning students']
+      f[label = 'Returning Student list']
+      n[label = 'rnorm@^{3}', shape = 'cylinder']
+      r[label = 'sample 0/1@^{4}', shape = 'cylinder']
+      s[label = '0 Credits']
+      t[label = 'Generate credit load']
+      
+      {rank = same; b, c}
+      {rank = same; e, f}
+      {rank = same; r, d}
+      {rank = same; n, t, s}
+      
+      #edges
+      b -> c
+      d -> {e, f}
+      c -> d
+      r -> d
+      e -> s
+      f -> t
+      n -> t
+      }
+      
+      subgraph cluster_1{
+      graph[shape = rectangle
+      label = 'Aggregate-level'
+      labelloc = t
+      fontsize = 25]
+      
+      node[shape = 'box']
+      g[label = 'Predict # of new students']
+      h[label = 'OLS@^{2}', shape = 'cylinder']
+      i[label = 'Sample credit load']
+      m[label = 'rnorm@^{3}', shape = 'cylinder']
+      o[label = 'Generate individual values']
+      v[label = 'Sample gender + credits@^{5}', shape = 'cylinder']
+      
+      {rank = same; g, h}
+      {rank = same; i, m}
+      {rank = same; v, o}
+      
+      #edges
+      m -> i
+      h -> g
+      g -> o
+      o -> i
+      v -> o
+      }
+      
+      subgraph cluster_2{
+      graph[label = 'Footnotes'
+      labelloc = t
+      penwidth = 0
+      ranksep = 0.02]
+      
+      node[shape = 'none']
+      x[label = '1: logit(y) = \u03b2*Gender + \u03b2*Cumulative credits + c(function of GDP) + \u03b5', x = 1, y = 1]
[…]

### 2021-06-06 Data Generation/Data_Generation.R

@@ -1,0 +1,282 @@
+library('truncnorm')
+library('rlist')
+
+# Initialize variables ----------------------------------------------------
+
+set.seed(100)
+
+#values for GDP. Random walk process
+periods <- 100
+GDP <- rnorm(1, mean = 0, sd = 1)
+
+for(i in 2:periods) {
+  GDP[i] <- GDP[i - 1] + rnorm(1, mean = 0, sd = 1)
+}
+
+#creating a data frame with semester binaries
+semester_variables <- data.frame(cbind(GDP,
+                                       rep_len(c(1, 0, 0),
+                                               length.out = periods),
+                                       rep_len(c(0, 1, 0),
+                                               length.out = periods),
+                                       rep_len(c(0, 0, 1),
+                                               length.out = periods)))
+
+colnames(semester_variables) <- c('GDP', 'Spring', 'Summer', 'Fall')
+
+
+#initial values for new student aggregate model
+B_GDP <- 2
+mean_new_students <- 150
+sd_new_students <- 25
+B_spring <- 3
+B_summer <- 0
+B_fall <- 6
+
+#initial values for t = 0 individual student creation
+proportion_female <- 0.5
+initial_students <- 900
+
+#parameters for assigning credit load
+mean_credits <- 9
+sd_credits <- 6
+
+#values for 'return' variable calculation
+B_Gender <- 0.1
+B_Credits <- 0.02
+C_constant <- 0.9
+
+# Aggregate prediction ----------------------------------------------------
+
+#general linear model for enrollment which has linear relationship with GDP,
+# semester fixed effects, and noise
+new_student_count <- semester_variables$GDP * B_GDP +
+  semester_variables$Spring * B_spring + 
+  semester_variables$Summer * B_summer +
+  semester_variables$Fall * B_fall + 
+  rnorm(nrow(semester_variables),
+        mean = 0,
+        sd = 1)
+
+#un-standardizing the calculation so we can get sensible numbers
+new_student_count <- round(new_student_count * sd_new_students + mean_new_students)
+
+
+# Create new students at t = 0 --------------------------------------------
+
+#create t = 0, initial data frame
+enrolled_students <- data.frame(matrix(c(1:initial_students,
+                                         #randomly assign gender
+                                         sample(c(0, 1),
+                                                size = initial_students,
+                                                replace = TRUE,
+                                                #probability of gender can be changed if needed
+                                                prob = c(proportion_female, 1 - proportion_female)),
+                                         #sample credits taken in the GIVEN SEMESTER. This part 
[…]

### 2021-06-06 Data_Generation.R

@@ -1,282 +1,0 @@
-library('truncnorm')
-library('rlist')
-
-# Initialize variables ----------------------------------------------------
-
-set.seed(100)
-
-#values for GDP. Random walk process
-periods <- 100
-GDP <- rnorm(1, mean = 0, sd = 1)
-
-for(i in 2:periods) {
-  GDP[i] <- GDP[i - 1] + rnorm(1, mean = 0, sd = 1)
-}
-
-#creating a data frame with semester binaries
-semester_variables <- data.frame(cbind(GDP,
-                                       rep_len(c(1, 0, 0),
-                                               length.out = periods),
-                                       rep_len(c(0, 1, 0),
-                                               length.out = periods),
-                                       rep_len(c(0, 0, 1),
-                                               length.out = periods)))
-
-colnames(semester_variables) <- c('GDP', 'Spring', 'Summer', 'Fall')
-
-
-#initial values for new student aggregate model
-B_GDP <- 2
-mean_new_students <- 150
-sd_new_students <- 25
-B_spring <- 3
-B_summer <- 0
-B_fall <- 6
-
-#initial values for t = 0 individual student creation
-proportion_female <- 0.5
-initial_students <- 900
-
-#parameters for assigning credit load
-mean_credits <- 9
-sd_credits <- 6
-
-#values for 'return' variable calculation
-B_Gender <- 0.1
-B_Credits <- 0.02
-C_constant <- 0.9
-
-# Aggregate prediction ----------------------------------------------------
-
-#general linear model for enrollment which has linear relationship with GDP,
-# semester fixed effects, and noise
-new_student_count <- semester_variables$GDP * B_GDP +
-  semester_variables$Spring * B_spring + 
-  semester_variables$Summer * B_summer +
-  semester_variables$Fall * B_fall + 
-  rnorm(nrow(semester_variables),
-        mean = 0,
-        sd = 1)
-
-#un-standardizing the calculation so we can get sensible numbers
-new_student_count <- round(new_student_count * sd_new_students + mean_new_students)
-
-
-# Create new students at t = 0 --------------------------------------------
-
-#create t = 0, initial data frame
-enrolled_students <- data.frame(matrix(c(1:initial_students,
-                                         #randomly assign gender
-                                         sample(c(0, 1),
-                                                size = initial_students,
-                                                replace = TRUE,
-                                                #probability of gender can be changed if needed
-                                                prob = c(proportion_female, 1 - proportion_female)),
-                                         #sample credits taken in the GIVEN SEMESTER. This part 
[…]

### 2021-06-06 Model_Building.R

@@ -1,116 +1,0 @@
-library('forecast')
-
-data <- read.csv('data.csv',
-                 colClasses = c('NULL', rep(NA, 6)))
-
-
-# Aggregate-only model ----------------------------------------------------
-
-#fit models to 1:t-1 data and predict t
-for(i in 5:nrow(data)) {
-  #fit simple lm model with semester and GDP terms
-  lm <- lm(Total_enrollment ~ GDP + Spring + Summer + Fall,
-           data = data[1:(i - 1),])
-  
-  #linear t + 1 prediction
-  data$lm_prediction[i] <- predict(lm,
-                                   data[i,!names(data) %in% 'Total_enrollment'])
-  
-  #find optimal arima model to 1:t-1 data
-  arima <- auto.arima(data$Total_enrollment[1:(i - 1)],
-                      allowdrift = FALSE)
-  
-  #arima t + 1 prediction
-  data$arima_prediction[i] <- predict(arima,
-                                      n.ahead = 1)$pred[1]
-  
-  #get rid of temporary objects
-  remove(lm)
-  remove(arima)
-}
-
-#calculate residuals for each type of model
-data$lm_residuals <- data$Total_enrollment - data$lm_prediction
-
-data$arima_residuals <- data$Total_enrollment - data$arima_prediction
-
-#calculate absolute percent error for each type of model
-data$lm_ape <- abs(data$lm_residuals) / data$Total_enrollment
-
-data$arima_ape <- abs(data$arima_residuals) / data$Total_enrollment
-
-# Stacked model -----------------------------------------------------------
-
-#read in student level data and keep only necessary columns
-student_data <- read.csv('student_level.csv')
-
-student_data <- student_data[, names(student_data) %in% c('returned', 'Gender', 'Cumulative_credits', 'semester')]
-
-#read in # of new students and coerce to vector for easier
-#time series modelling
-new_student_count <- read.csv('new_student_count.csv',
-                              colClasses = c('NULL', NA))
-
-new_student_count <- as.vector(new_student_count$new_student_count)
-
-#set number of iterations
-semesters <- length(unique(student_data$semester))
-
-#create data frame to hold predictions
-stacked_model_predictions <- data.frame(matrix(nrow = semesters,
-                                               ncol = 3))
-
-colnames(stacked_model_predictions) <- c('Retained_prediction', 'New_prediction', 'Combined_prediction')
-
-for(i in 4:(semesters - 1)) {
-  #train dataset = all observations where semester <= i
[…]

### 2021-06-06 Modeling/Model_Building.R

@@ -1,0 +1,116 @@
+library('forecast')
+
+data <- read.csv('data.csv',
+                 colClasses = c('NULL', rep(NA, 6)))
+
+
+# Aggregate-only model ----------------------------------------------------
+
+#fit models to 1:t-1 data and predict t
+for(i in 5:nrow(data)) {
+  #fit simple lm model with semester and GDP terms
+  lm <- lm(Total_enrollment ~ GDP + Spring + Summer + Fall,
+           data = data[1:(i - 1),])
+  
+  #linear t + 1 prediction
+  data$lm_prediction[i] <- predict(lm,
+                                   data[i,!names(data) %in% 'Total_enrollment'])
+  
+  #find optimal arima model to 1:t-1 data
+  arima <- auto.arima(data$Total_enrollment[1:(i - 1)],
+                      allowdrift = FALSE)
+  
+  #arima t + 1 prediction
+  data$arima_prediction[i] <- predict(arima,
+                                      n.ahead = 1)$pred[1]
+  
+  #get rid of temporary objects
+  remove(lm)
+  remove(arima)
+}
+
+#calculate residuals for each type of model
+data$lm_residuals <- data$Total_enrollment - data$lm_prediction
+
+data$arima_residuals <- data$Total_enrollment - data$arima_prediction
+
+#calculate absolute percent error for each type of model
+data$lm_ape <- abs(data$lm_residuals) / data$Total_enrollment
+
+data$arima_ape <- abs(data$arima_residuals) / data$Total_enrollment
+
+# Stacked model -----------------------------------------------------------
+
+#read in student level data and keep only necessary columns
+student_data <- read.csv('student_level.csv')
+
+student_data <- student_data[, names(student_data) %in% c('returned', 'Gender', 'Cumulative_credits', 'semester')]
+
+#read in # of new students and coerce to vector for easier
+#time series modelling
+new_student_count <- read.csv('new_student_count.csv',
+                              colClasses = c('NULL', NA))
+
+new_student_count <- as.vector(new_student_count$new_student_count)
+
+#set number of iterations
+semesters <- length(unique(student_data$semester))
+
+#create data frame to hold predictions
+stacked_model_predictions <- data.frame(matrix(nrow = semesters,
+                                               ncol = 3))
+
+colnames(stacked_model_predictions) <- c('Retained_prediction', 'New_prediction', 'Combined_prediction')
+
+for(i in 4:(semesters - 1)) {
+  #train dataset = all observations where semester <= i
[…]

### 2021-06-06 Paper/Paper_Draft.Rmd
060621 updates

@@
 output:
+  html_document: default
   word_document: default
-  html_document: default
 bibliography: Citations.bib
@@
 Finally, the individual level predictions were aggregated and summed with the number of new enrollees to produce total headcount for a given semester.
-This process is visualized in chart 1.
+This process is visualized below.
+Nodes in yellow are 'outputs' of the generation process which are used in the model fitting section.
 
 ```{r, echo = FALSE}
-knitr::include_graphics('Data Generation Process.png')
+library('DiagrammeR')
+
+grViz("
+digraph data_generation {
+
+  graph[rankdir = TD]
+
+      subgraph cluster_0{
+
+      graph[shape = rectangle
+      label = 'Individual-level'
+      labelloc = t
+      fontsize = 25
+      rankdir = TD]
+
+      node[shape = 'box']
+      a[label = 'Students @ t-1']
+      b[label = 'Logistic\nregression', shape = 'cylinder']
+      c[label = 'Predict probability of return']
+      d[label = 'Generate return status']
+      e[label = 'Non-returning students']
+      f[label = 'Returning Student']
+      n[label = 'Random sample', shape = 'cylinder']
+      r[label = 'Random sample', shape = 'cylinder']
+      t[label = 'Credit load']
+
+      {rank = same; b, c}
+      {rank = same; e, f}
+      {rank = same; r, d}
+      {rank = same; n, t}
+
+      #edges
+      a -> c
+      b -> c
+      d -> {e, f}
+      c -> d
+      r -> d
+      f -> t
+      n -> t
+      c -> r[style = dashed]
+      }
+
+      subgraph cluster_1{
+      graph[shape = rectangle
+      label = 'Aggregate-level'
+      labelloc = t
+      fontsize = 25]
+
+      node[shape = 'box']
+      g[label = 'Number of New Students']
+      ab[label = 'Create N individual records']
+      h[label = 'OLS', shape = 'cylinder']
+      aa[label = 'De-normalize & round', shape = 'cylinder']
+      o[label = 'Individual variable values']
+      ac[label = 'Random sample', shape = 'cylinder']
+
+      {rank = same; g, h, aa}
+      {rank = same; o, ac}
+
+      #edges
+      h -> aa
+      aa -> g
+      g -> ab
+      ab -> o
+      ac -> o
+      }
+
+      node[shape = 'box']
+      j[label = 'Join']
+      l[label = 'New students']
+      q[label = 'Individual records @ t']
+      u[label = 'Returning students']
+      ad[label = 'Aggregate']
+      ae[label = 'Total enrollment @ t']
+
+      #edges
+      o -> l
+      {l, u} -> j
+      j -> q
+      t -> u
+      j -> ad
+      ad -> ae
+
+}")
 ```
@@
 To demonstrate the value in building theory-informed predictive models, I fit and compare several such models.
-I test autoregressive integrated moving average (ARIMA) and simple linear models at the aggregate level, as these are the most common approaches in the extent literature and compare them to three stacked models.
-For the first stacked model, I predict the number of returning students at the individual level using logistic regression and use that number as a linear predictor of the *total* number of enrolling students at semester t + 1.
-Second, I fit a theoretically misspecified model, predicting the number of returning students at the individual level using *only* lower order terms in a logistic regression and the number of new students using ARIMA.
-Finally, I fit a correctly specified model which differs only insofar as it uses a correct higher order term in the individual prediction section.
+I test autoregressive integrated moving average (ARIMA) and simple linear models at the aggregate level, as these are the most common approaches in the extent literature^[Exponential smoothing approaches are also common early in the enrollment forecasting literature; X, for instance, lists such an approach as one of the most common ways community college practitioners approach modeling.
+However, this approach sees little use - at least in its simple form - among contemporary practitioners who prefer other methods of time series modeling.
+For this reason, I elect not to include an exponential smoothing model in my comparison] and compare them to two stacked models.
+For the first stacked model, I fit a theoretically misspecified model, predicting the number of returning students at the individual level using  non-polynomial terms in a logistic regression and the number of new students using ARIMA.
+For the second, I fit a correctly specified model with the correct polynomial term.
 
-I compare the results of these models using the Mean Absolute Percent Error (MAPE), Mean Absolute Standard Error (MASE), and Root Mean Squared Error (RMSE), three widely used measures of model accuracy.
+I compare the results of these models using the Mean Mean Absolute Error (MASE), Absolute Percent Error (MAPE), and Root Mean Squared Error (RMSE), three widely used measures of model accuracy.
 
 # Results
+
+```{r, echo = FALSE}
+library('ggplot2')
+library('reshape2')
+
+predictions <- read.csv('predictions.csv')
+
+predictions <- predictions[,-1]
+
+predictions <- predictions[,c(5, grep('prediction', colnames(predictions)))]
+
+predictions$semester <- 1:nrow(predictions)
+
+predictions <- melt(predictions,
+                    id.vars = 'semester')
+
+predictions <- predictions[complete.cases(predictions),]
+
+predictions <- predictions[-which(predictions$variable == 'stacked_predictions_1'),]
+
+ggplot(data = predictions) +
+  geom_line(aes(x = semester,
+                y = value,
+                color = variable,
+                group = variable)) +
+  theme()
+
+```
+
+```{r, echo = FALSE}
+diagnostics <- read.csv('diagnostics.csv')
+
+diagnostics <- diagnostics[,-1]
+
+knitr::kable(diagnostics)
+```
 
@@
 A skeptical reader may justifiably question the value in the above exercise.
-After all, of course the better specified model predicted more accurately, we *knew* the correct parameters before starting.
+After all, of course the better specified model predicted more accurately, we *knew* the true parameters and data generating process before starting.
 This is, however, exactly the point.
@@
 Employing stacked models, as shown above, can be a simple manner of doing so.
-
-# Future Research
 

### 2021-06-06 Data Generation/Data Generation Diagram.R

@@ -52,7 +52,7 @@ digraph data_generation {
       g[label = 'Predict # of new students']
       h[label = 'OLS@^{2}', shape = 'cylinder']
       i[label = 'Sample credit load']
-      m[label = 'rnorm@^{3}', shape = 'cylinder']
+      m[label = '@@1', shape = 'cylinder']
       o[label = 'Generate individual values']
       v[label = 'Sample gender + credits@^{5}', shape = 'cylinder']
       
@@ -107,4 +107,6 @@ digraph data_generation {
       m -> x[style = 'invis']
       
       {rank = same; a, w}
-}")
+}
+
+[1]: 'test'")

### 2021-06-06 Data Generation/Data_Generation.R

@@ -269,14 +269,17 @@ semester_variables$Total_credits_taken <- aggregate(temp$Credits,
                                                     FUN = sum)[,2]
 
 #write out aggregate file
-#write.csv(semester_variables,
-#          file = 'data.csv')
+write.csv(semester_variables,
+          file = 'data.csv')
 
 #write out individual file
-#write.csv(temp,
-#          file = 'student_level.csv')
+write.csv(temp,
+          file = 'student_level.csv')
+
+
+
 
 #write out new student count file
-#write.csv(data.frame(new_student_count),
-#          file = 'new_student_count.csv')
-
+write.csv(data.frame(cbind(new_student_count,
+                           GDP)),
+          file = 'new_student_count.csv')

### 2021-06-06 Modeling/Model_Building.R

@@ -1,8 +1,12 @@
 library('forecast')
+library('scales')
 
 data <- read.csv('data.csv',
                  colClasses = c('NULL', rep(NA, 6)))
 
+for(i in 1:nrow(data)) {
+  data$lag[i] <- data$Total_enrollment[i - 1]
+}
 
 # Aggregate-only model ----------------------------------------------------
 
@@ -15,6 +19,14 @@ for(i in 5:nrow(data)) {
   #linear t + 1 prediction
   data$lm_prediction[i] <- predict(lm,
                                    data[i,!names(data) %in% 'Total_enrollment'])
+  
+  #fit lagged linear model
+  lm_lagged <- lm(Total_enrollment ~ GDP + lag +Spring + Summer + Fall,
+                  data = data[1:(i - 1),])
+  
+  #linear with lat t + 1 prediction
+  data$lm_lagged_prediction[i] <- predict(lm_lagged,
+                                          data[i,!names(data) %in% 'Total_enrollment'])
   
   #find optimal arima model to 1:t-1 data
   arima <- auto.arima(data$Total_enrollment[1:(i - 1)],
@@ -32,10 +44,14 @@ for(i in 5:nrow(data)) {
 #calculate residuals for each type of model
 data$lm_residuals <- data$Total_enrollment - data$lm_prediction
 
+data$lm_lagged_residuals <- data$Total_enrollment - data$lm_lagged_prediction
+
 data$arima_residuals <- data$Total_enrollment - data$arima_prediction
 
 #calculate absolute percent error for each type of model
 data$lm_ape <- abs(data$lm_residuals) / data$Total_enrollment
+
+data$lm_lagged_ape <- abs(data$lm_lagged_residuals) / data$Total_enrollment
 
 data$arima_ape <- abs(data$arima_residuals) / data$Total_enrollment
 
@@ -66,20 +82,33 @@ for(i in 4:(semesters - 1)) {
   #train dataset = all observations where semester <= i
   train <- student_data[which(student_data$semester <= i),]
   
-  #fit model to training data
+  #fit stacked model 1 (no polynomial term) to training data
   individual_model <- glm(returned ~ Gender + Cumulative_credits,
-                          data = train)
+                          data = train,
+                          family = 'binomial')
+  
+  #fit stacked model 2 (polynomial term)
+  individual_model_1 <- glm(returned ~ Gender + poly(Cumulative_credits, 2) + Cumulative_credits,
+                            data = train,
+                            family = 'binomial')
   
   #prediction = t + 1
[…]

### 2021-06-07 Paper/Paper_Draft.Rmd
060621 update

@@
 
-# Methods
+# Method
 
@@
 
-I compare the results of these models using the Mean Mean Absolute Error (MASE), Absolute Percent Error (MAPE), and Root Mean Squared Error (RMSE), three widely used measures of model accuracy.
+I compare the results of these models using the Mean Absolute Error (MASE), Mean Absolute Percent Error (MAPE), and Root Mean Squared Error (RMSE), three widely used measures of predictive accuracy.
+To ensure results are not the product of random chance, I fit each model to 1000 unique datasets and present the mean value for each measure of model accuracy.
 
 # Results
+
+While all tested models predict enrollment with some degree of accuracy, there are large differences between models.
+Simple linear models without lag terms perform poorest, with MAPE scores near 9%.
+Simple ARIMA models do not fare much better with MAPE scores hovering near 6%.
+There are smaller differences between simple linear models with lag terms and stacked models.
+However, stacked models, both with and without higher order terms, outperform the former, if only by a few tenths of a percent in MAPE.
+
+Though the differences between stacked and linear lagged models are statistically small, their substantive importance should not be dismissed.
+Such differences can easily mean missing predictions by several hundred students at a community college with a student body in the tens of thousands.
+Moreover, it should be noted that, while the linear model with a lagged term is more parsimonious than the stacked models, parsimony for its own sake is dubious in circumstances where predictive accuracy is paramount
 
@@
 Practitioners ought to take pains to incorporate theoretical knowledge in predictive forecasts.
-Employing stacked models, as shown above, can be a simple manner of doing so.
+Employing stacked models, as shown above, can be a simple and effective manner of doing so.
 

### 2021-06-07 Data Generation/Data_Generation.R

@@ -3,7 +3,7 @@ library('rlist')
 
 # Initialize variables ----------------------------------------------------
 
-set.seed(100)
+set.seed(1000)
 
 #values for GDP. Random walk process
 periods <- 100
@@ -12,6 +12,8 @@ GDP <- rnorm(1, mean = 0, sd = 1)
 for(i in 2:periods) {
   GDP[i] <- GDP[i - 1] + rnorm(1, mean = 0, sd = 1)
 }
+
+GDP <- GDP + abs(min(GDP))
 
 #creating a data frame with semester binaries
 semester_variables <- data.frame(cbind(GDP,
@@ -60,7 +62,8 @@ new_student_count <- semester_variables$GDP * B_GDP +
 
 #un-standardizing the calculation so we can get sensible numbers
 new_student_count <- round(new_student_count * sd_new_students + mean_new_students)
-
+new_student_count
+plot(GDP)
 
 # Create new students at t = 0 --------------------------------------------
 
@@ -275,8 +278,6 @@ write.csv(semester_variables,
 #write out individual file
 write.csv(temp,
           file = 'student_level.csv')
-
-
 
 
 #write out new student count file

### 2021-06-07 Modeling/Model_Building.R

@@ -105,6 +105,14 @@ for(i in 4:(semesters - 1)) {
                                                                          test, type = 'response')) * nrow(test)
   
   
+  #build linear model
+  #lm <- lm(new_student_count[1:i] ~ GDP + Spring + Summer,
+  #         data = data[1:i,])
+  
+  #predict linear model
+  #stacked_model_predictions$New_prediction[i + 1] <- predict(lm,
+  #                                                           data[i + 1,])
+  
   #build arima model for new students only. As above, time through i
   arima <- auto.arima(new_student_count[1:i],
                       allowdrift = FALSE,
@@ -189,3 +197,6 @@ write.csv(diagnostics,
 #write out predictions to plot for paper
 write.csv(data,
           file = 'predictions.csv')
+
+stacked_model_predictions
+summary(lm)

### 2021-06-07 Paper/Paper_Draft.Rmd
060721 update

--- 2021-06-06 Paper/Paper_Draft.Rmd
@@
 
+```{r, setup, echo = FALSE}
+library('ggplot2')
+library('reshape2')
+library('DiagrammeR')
+library('knitr')
+library('scales')
+```
+
 # Introduction
 
-Enrollment forecasting is one of the most consequential tasks a given Institutional Research department will undertake.
-The pressures of getting it 'right' are especially pronounced in the community college setting.
-As open enrollment institutions that are uniquely reliant on public funding, accurate enrollment forecasts are critical for guaranteeing adequate state-level funding and proper planning.
+Enrollment forecasting is one of the most consequential tasks Institutional Research departments undertake.
+As open enrollment institutions that are uniquely reliant on public funding, community colleges need accurate enrollment forecasts for guaranteeing adequate state-level funding and proper planning.
 Moreover, community colleges operate under unique data availability constraints and admission processes that make the task difficult.
@@
 Despite the importance of this task, the extant literature on enrollment forecasting - especially in the community college context - is slim and relatively disconnected from theoretical advances in the study of the predictors of retention and enrollment.
-This manifests itself in two main problems.
+This manifests itself in two main problems:
+
+1. There is little literature that speaks to this problem in the context of community colleges
+2. Predictive models rarely reflect theoretical knowledge
 
@@
 [@aksenova_enrollment_2006], [@chen_integrated_2008], [@nandeshwar_enrollment_2009], and [@slim_predicting_2018] for instance, exclusively apply their models to the selective enrollment context.
-Indeed, X's models are built to specifically predict the likelihood of enrollment *after* acceptance; a process that has no direct analogue in the open enrollment context.
 Work addressing the problem in the community college context is largely dated.
@@
 [@pennington_community_2002] represents the most recent and comprehensive project addressing this issue.
-Numerous dissertations (see: x, y, and z) have reproduced and expanded this work in varying local contexts, adding important contextual knowledge but few explicitly addressing modeling choices.
 In all, the literature that speaks to predictive modeling of enrollment in the community college context is limited.
@@
 Second, predictive models of enrollment do not reflect our theoretical understanding of the processes that are being modeled.
-Much of the modeling in the extant literature falls into one of two camps: time series approaches at the aggregate level (ARIMA) or linear modeling techniques (OLS, SVG, etc.).
+Much of the modeling in the extant literature falls into one of two camps: time series approaches at the aggregate level (ARIMA) or linear modeling techniques (OLS).
 While such approaches produce perfectly acceptable predictions, such models are rarely reflective of the theoretical processes they attempt to capture.
 [@chen_integrated_2008], for instance, compares predictions from both, ARIMA and linear models though uses the aggregate number of enrollments in a given semester as the target variable despite noting that '[...] the portion of a university’s student enrollees (freshmen) depends on the number of high school graduates within the state [...] [and] the same pool of eligible enrolled or returning students (one-year lagged OSU enrollment).' The recognition that two distinct sub-populations constitute the population of enrollees is not reflected in the modeling which treats each semester's cohort uniformly^[See also [@pfitzner_shortterm_1987] for an early implementation of time series modeling in the community college context that approaches the problem at the aggregate level].
-By contrast, [@trusheim_predictive_2011] does model returning and new enrollees separately, and while this serves as a good example of using theoretical knowledge as a foundation for in empirical exercises, the methods used to achieve predictions are fairly unsophisticated.
+By contrast, [@trusheim_predictive_2011] does model returning and new enrollees separately, and serves as a good example of using theoretical knowledge as a foundation for in empirical exercises, though the methods used to achieve predictions are fairly unsophisticated.
 
-Here, I argue that, in the interest of predictive accuracy and knowledge accumulation, our statistical models ought to be theoretically informed.
+Here, I argue that, in the interest of predictive accuracy, our statistical models ought to be theoretically informed.
 To demonstrate this, I present a hypothetical example with generated data, an approach widely used in the study of statistics.
@@
 1. Simulated data are generated with *known* parameters, allowing us to determine how accurately models recover *true* estimates
-2. Simulated data allow for generalizability. Data generating processes may differ dramatically across institutions.
-3. Simulated data allow us to demonstrate theoretically important concepts by assuring these concepts are the *only* relevant ones in our model.
+2. Simulated data allow for generalizability. Enrollment processes may differ dramatically across institutions.
+3. Simulated data allow us to demonstrate theoretically important concepts by ensuring these concepts are the *only* relevant ones in our model.
 
@@
 This process is visualized below.
-Nodes in yellow are 'outputs' of the generation process which are used in the model fitting section.
+Nodes in yellow are 'outputs' of the generative  process which are used in the model fitting section.
 
 ```{r, echo = FALSE}
-library('DiagrammeR')
 
@@
       l[label = 'New students']
-      q[label = 'Individual records @ t']
+      q[label = 'Individual records @ t', style = 'filled', fillcolor = yellow]
       u[label = 'Returning students']
       ad[label = 'Aggregate']
-      ae[label = 'Total enrollment @ t']
+      ae[label = 'Total new enrollment @ t', style = 'filled', fillcolor = yellow]
 
@@
       t -> u
-      j -> ad
+      l -> ad
       ad -> ae
@@
 The formula used to generate each student's probability of returning is given below; the coefficients associated with each term as well as the values for the constant and error parameters can be found in the Appendix.
-As noted above, I chose to generate data using a simple model for the purpose of demonstration; Gender and cumulative credit load are widely cited predictors of retention.
+As noted above, I chose to generate data using a simple model for the purpose of demonstration; Gender and cumulative credit load are not the only predictors of retention but are widely cited and important variables.
 
@@
 
-The number of new enrollees in a given semester was generated at the aggregate level and is given in heuristic form below.
+The number of new enrollees in a given semester was generated at the aggregate level and is given below.
 The number of new enrollees in a given semester is a linear function of change in GDP and semester such that higher GDP growth is associated with greater enrollment.
@@
 
-# Methods
+# Method
 
-Modeling choices ought to reflect our underlying knowledge about the data generating process.
+Modeling choices ought to reflect our knowledge about the underlying data generating process.
 In this example, we know that the processes that generated our observations vary across sub-populations.
@@
 
-I compare the results of these models using the Mean Mean Absolute Error (MASE), Absolute Percent Error (MAPE), and Root Mean Squared Error (RMSE), three widely used measures of model accuracy.
+I compare the results of these models using the Mean Absolute Error (MAE), Mean Absolute Percent Error (MAPE), and Root Mean Squared Error (RMSE), three widely used measures of predictive accuracy.
+To ensure results are not the product of random chance, I fit each model to 100 uniquely generated data sets and present the mean value and standard error for each measure of model accuracy.
 
@@
 
+While all tested models predict enrollment with some degree of accuracy, there are large differences between models.
+Simple linear models perform significantly worse across all measures of predictive accuracy, with large mean MAE, MAPE, and RMSE scores.
+Simple ARIMA models do not fare much better though represent an improvement over simple linear models.
+
 ```{r, echo = FALSE}
-library('ggplot2')
-library('reshape2')
 
-predictions <- read.csv('predictions.csv')
+temp <- read.csv('aggregate_metrics.csv')
 
-predictions <- predictions[,-1]
+#calculate standard error for each of my diagnostic
+#measure means
+st.error <- function(x) sd(x) / sqrt(length(x))
 
-predictions <- predictions[,c(5, grep('prediction', colnames(predictions)))]
+table <- data.frame(
+  cbind(
+    #mean MAE
+    aggregate(temp$MAE,
+              by = list(temp$Model),
+              FUN = mean),
+    #standard error MAE
+    aggregate(temp$MAE,
+              by = list(temp$Model),
+              FUN = st.error)[,2],
+    #mean MAPE
+    aggregate(temp$MAPE,
+              by = list(temp$Model),
+              FUN = mean)[,2],
+    #standard error MAPE
+    aggregate(temp$MAPE,
+              by = list(temp$Model),
+              FUN = st.error)[,2],
+    #mean RMSE
+    aggregate(temp$RMSE,
+              by = list(temp$Model),
+              FUN = mean)[,2],
+    #standard error RMSE
+    aggregate(temp$RMSE,
+              by = list(temp$Model),
+              FUN = st.error)[,2]
+  ))
 
-predictions$semester <- 1:nrow(predictions)
+colnames(table) <- c('Model', 
+                     'MAE Mean', 'MAE St.error',
+                     'MAPE Mean', 'MAPE St.error',
+                     'RMSE Mean', 'RMSE St.error')
 
-predictions <- melt(predictions,
-                    id.vars = 'semester')
+table[,c(2, 3, 6, 7)] <-  round(table[,c(2, 3, 6, 7)], 2)
 
-predictions <- predictions[complete.cases(predictions),]
+table[,4] <- percent(table[,4])
 
-predictions <- predictions[-which(predictions$variable == 'stacked_predictions_1'),]
+table[,5] <- percent(table[,5], 0.001)
 
-ggplot(data = predictions) +
-  geom_line(aes(x = semester,
-                y = value,
-                color = variable,
-                group = variable)) +
-  theme()
+kable(table)
 
@@
 
-```{r, echo = FALSE}
-diagnostics <- read.csv('diagnostics.csv')
+Comparisons between linear models with lag terms and stacked models illustrates the critical point of this exercise.
+While linear models with lag terms perform essentially on par with the stacked models in terms of MAE and MAPE, linear models perform significantly worse on RMSE and are more volatile in this regard.
+In essence, when linear models with lag terms get it wrong, they get it much 'wronger' than stacked models.
+Additionally, across datasets, the prevalence of highly 'wrong' values varies considerably with linear models that include lag terms; hence its volatility.
+Only stacked models predict enrollment precisely and accurately.
 
-diagnostics <- diagnostics[,-1]
+```{r, echo = FALSE, message = FALSE}
+temp <- read.csv('aggregate_metrics.csv')
 
-knitr::kable(diagnostics)
+temp <- temp[,-1]
+
+temp$index <- 1:nrow(temp)
+
+viz <- melt(temp,
+            id.vars = c('index', 'Model'))
+
+ggplot(data = viz) +
+  geom_histogram(aes(x = value)) +
+  facet_grid(rows = vars(viz$Model),
+             cols = vars(viz$variable),
+             scales = 'free_x')
+
 ```
+
+Finally, it should be noted that the correctly specified stacked model does not outperform the parsimonious stacked model though differences are statistically negligible across all measures of predictive accuracy.
+Neither are volatile in the same way that linear models with lag terms are.
 
@@
 
-# Appendix B: Cut Parts
-
-In this paper, I have demonstrated the value in bringing theoretical knowledge to bear on predictive models of enrollment.
-Community colleges construct such models under unique constraints.
-Moreover, the financial consequences of inaccurate models is more acutely felt than at other types of institutions of higher education.
-While the extant literature has sought improvements to predictive accuracy by using 'brute-force' methods - e.g.
-increasing the number of features in models, implementing more model types, exhaustively searching across model hyper-parameters - I argue that these efforts essentially re-invent the wheel insofar as they 'rediscover' relationships that are already well established.
-Our theoretical knowledge of the predictors of enrollment and persistence is robust.
-That knowledge should be reflected in our empirical models.
-
-# Appendix C: Summary Statistics
+# Works Cited

### 2021-06-07 Robustness Checks/Data_Generation.R

@@ -1,0 +1,514 @@
+library('truncnorm')
+library('rlist')
+library('forecast')
+library('scales')
+library('ggplot2')
+library('reshape2')
+
+#everything in this file is just a combination of the data generation and
+#modeling files up to the final section which shows how I generated the
+#aggregate fit values
+
+# Initialize variables ----------------------------------------------------
+
+set.seed(1000)
+
+#values for GDP. Random walk process
+periods <- 100
+GDP <- rnorm(1, mean = 0, sd = 1)
+
+for(i in 2:periods) {
+  GDP[i] <- GDP[i - 1] + rnorm(1, mean = 0, sd = 1)
+}
+
+GDP <- GDP + abs(min(GDP))
+
+#creating a data frame with semester binaries
+semester_variables <- data.frame(cbind(GDP,
+                                       rep_len(c(1, 0, 0),
+                                               length.out = periods),
+                                       rep_len(c(0, 1, 0),
+                                               length.out = periods),
+                                       rep_len(c(0, 0, 1),
+                                               length.out = periods)))
+
+colnames(semester_variables) <- c('GDP', 'Spring', 'Summer', 'Fall')
+
+
+#initial values for new student aggregate model
+B_GDP <- 2
+mean_new_students <- 150
+sd_new_students <- 25
+B_spring <- 3
+B_summer <- 0
+B_fall <- 6
+
+#initial values for t = 0 individual student creation
+proportion_female <- 0.5
+initial_students <- 900
+
+#parameters for assigning credit load
+mean_credits <- 9
+sd_credits <- 6
+
+#values for 'return' variable calculation
+B_Gender <- 0.1
+B_Credits <- 0.02
+C_constant <- 0.9
+
+# Diagnosis and robustness checks -----------------------------------------
+
+
+#enrollment seems steady over time so it looks like
+#the coefficients and constants specified are ok.
+#trying to keep it around a thousand
+plot(aggregate(temp$likelihood_of_return,
+               by = list(temp$semester),
+               FUN = length))
+
+#credit load per semester also seems robust over time
+#There's some variation but nothing that's concerning
+aggregate(temp$Credits,
+          by = list(temp$semester),
+          FUN = sum)
+
+#our initial specification of the cumulative credit load
+#seems too high. However, it looks like we hit an equilibrium
+#really fast. It seems that the penalty coefficient of
+#credits -> likelihood of return is doing its job. Even so,
[…]
